---
title: "Cedar Creek data analysis"
author: "Melissa DeSiervo" mhdesiervo@gmail.com; mdesierv@uwyo.edu
date: "3/25/2022"
output: html_document
---


library(plyr)
library(dplyr)
library(ggplot2)
library(ape)
library(devtools)
library(tidyr)
library(vegan) ##must be the most recent version from github##
library(viridis)
library(stringr)
library(here)
library(tidyverse)
library(tsvr)
library(ggpubr)
library(ecotraj)
library(investr)
library(data.table)
library(trajr)
library(gtable)
library(gridExtra)
library(codyn)
library(indicspecies)
library(BiodiversityR)




```{r aictable function}

aictable<-function(X,m){
  #
  #  This function will create a full model selection table based on AICc.
  #  
  #  Inputs:
  #
  #  X is the AIC table produced by R by using AIC(model1,model2, ...)
  #  m is the sample size
  #
  #
  rnames<-row.names(X)
  AICc<-X$AIC+2*X$df*(X$df+1)/(m-X$df-1)     #small-sample correction
  logL<-X$df-X$AIC/2                         #Log-likelihood
  tab<-data.frame(X[,1],logL,AICc)           #Remove AIC column; add logL and AICc
  colnames(tab)[1]<-c("Params")              #Rename "df" column   
  row.names(tab)<-rnames
  tab<-tab[order(tab$AICc),]                 #Sort by ascending AICc value
  deltaAICc<-tab$AICc-min(tab$AICc)          #Delta AICc
  weight<-exp(-deltaAICc/2)/sum(exp(-deltaAICc/2))  #Weights
  cumwt<-weight                              #Column for cumulative weight
  for(i in 2:dim(X)[1]){                  
    cumwt[i]<-cumwt[i-1]+cumwt[i]              #Accumulate weight from the top
  }
  tab<-data.frame(tab,deltaAICc,weight,cumwt)
  tab<-round(tab,4)
  tab
}

```


```{r upload and clean datasets}


# Reading in and cleaning Cedar Creek data###

#Read in e001 data## Plot data for the unplowed fields##

d1s <- read.csv(here("data/e001-soil-cn-2019-09-13.csv"), header=T, strip.white=T, skip=1)

# Delete some extra variables and add dummies to stack exp1 & exp2 data
d1s$natm.nadd <- NULL
d1s$fertadd <- NULL
d1s$burn.trt.1992 <- NA
d1s$ntrt.last.year <- 'CurrentYear'
d1s$ntrt.origin <- 1 # No cessation in E001
d1s$disk <- 0 # E001 was not disked before experiment (disturbed vegetation)
d1s$exp <- 1

#Read in e002 data ##Plot data for the plowed fields##

d2s <- read.csv(here("data/e002-soil-cn-2019-09-13.csv"), header=T, strip.white=T, skip=1)

## Delete some extra variables and add dummies to stack exp1 & exp2 data
d2s$X <- NULL
d2s$date <- NULL
d2s$BurnRecBeginning1992 <- NULL
d2s$NtrtRecBefore1992 <- NULL
d2s$TrtRecAfter1991 <- NULL
d2s$TrtRecAfter2013 <- NULL
d2s$burn <- NA
d2s$fence <- NA
d2s$fence.origin <- NA
d2s$depth <- NA
d2s$disk <- 1 # E002 was disked before experiment (disturbed vegetation)
d2s$exp <- 2


#Additioanl plot info for both experiments##

d3s <- read.csv(here("data/E001-E002-Soil-CN-2018.csv"), header=T, strip.white=T, skip=0, na.strings=c(""))


# Delete samples that have errors
d3s <- d3s[is.na(d3s$FLAGGED.FOR.RERUN..S.Spillage..E.Error.),]

# Delete some extra variables and add dummies to stack exp1 & exp2 data
d3s$FLAGGED.FOR.RERUN..S.Spillage..E.Error. <- NULL
d3s$Sample <- NULL
d3s$X <- NULL
d3s$burn <- NA
d3s$burn.origin <- NA
d3s$burn.trt.1992 <- NA
d3s$fence <- NA
d3s$fence.origin <- NA
d3s$depth <- NA
d3s$disk <- NA
d3s$nadd <- NA
d3s$ntrt <- NA
d3s$ntrt.last.year <- NA
d3s$ntrt.origin <- NA
d3s$trt.origin <- NA  

d3s$disk[d3s$exp==1] <- 0 # E001 was not disked before experiment (intact vegetation)
d3s$disk[d3s$exp==2] <- 1 # E002 was disked before experiment (disturbed vegetation)

# Check that data sets have the same variables
sort(names(d1s))
sort(names(d2s))
sort(names(d3s))

# Stack exp1 & 2 data
ds <- rbind(d1s,d2s,d3s)


#Read in e001 aboveground plant biomass data ##

d1a <- read.csv(here("data/e001-aboveground-mass-2019-09-13.csv"), header=T, strip.white=T, skip=1)



# Delete some extra variables and add dummies to stack exp1 & exp2 data
d1a$natm.nadd <- NULL
d1a$fertadd <- NULL
d1a$burn.trt.1992 <- NA
d1a$ntrt.last.year <- "CurrentYear" # No cessation in E001
d1a$ntrt.origin <- 1 # No cessation in E001
d1a$subplot <- NA
d1a$disk <- 0 # E001 was not disked before experiment (disturbed vegetation)
d1a$exp <- 1


#Read in e002 aboveground plant biomass data ##
d2a <- read.csv(here("data/e002-aboveground-mass-2019-09-13.csv"), header=T, strip.white=T, skip=1)


# Delete some extra variables and add dummies to stack exp1 & exp2 data
d2a$X <- NULL
d2a$date <- NULL
d2a$burn <- NA
d2a$fence <- NA
d2a$fence.origin <- NA
d2a$disk <- 1 # E002 was disked before experiment (intact vegetation)
d2a$exp <- 2


# Check that data sets have the same variables
sort(names(d1a))
sort(names(d2a))


# Stack exp1 & 2 data
da <- rbind(d1a,d2a)

# Delete records with key missing data
da <- da[!is.na(da$mass),]

# Disking not done if field D
da <- da[da$field != 'D',]
da$field <- as.factor(as.character(da$field))


# If subplot is missing specify it as a whole plot
da$subplot[is.na(da$subplot)] <- "Whole"

# Code other Nutrient additions
da$other.add <- 1
da$other.add[da$ntrt == 9] <- 0

# Capitalize species to get rid of capilization differences in spelling
da$species <- toupper(as.character(da$species))


# Capitalize species to get rid of capilization differences in spelling
da$species <- toupper(as.character(da$species))

# Don't include woody plants in biomass as there are some big trees in samples
da$live <- 1
da$sorted <- 1
da$wood <- 0
da$vasc <- 1

# Do some general substitutions

da$species <- gsub("APOCYNUM CANNABINUM", "APOCYNUM ANDROSAEMIFOLIUM", da$species) ## MD 11/1 based off email with Eric##
da$species <- gsub("MISC. FORB", "MISCELLANEOUS FORB", da$species)
da$species <- gsub("SEDGES", "CAREX SP.", da$species)
da$species <- gsub("QUERCUS RUBRUM", "QUERCUS RUBRA", da$species)

# Find litter and code as not alive
sel<-da$species == 'MISCELLANEOUS LITTER'
da$live[sel] <- 0

sel<-grep("PINE", da$species)
da$live[sel] <- 0

# Code unsorted material as not being sorted
sel<-grep("MISCELLANEOUS", da$species)
da$sorted[sel] <- 0

sel<-grep("FUNGI", da$species)
da$sorted[sel] <- 0

sel<-grep("MOSS", da$species)
da$sorted[sel] <- 0

sel<-grep("LICHEN", da$species)
da$sorted[sel] <- 0

# Woody stuff
sel<-grep("ACER NEGUNDO", da$species)
da$wood[sel] <- 1
sel<-grep("CEANOTHUS", da$species)
da$wood[sel] <- 1
sel<-grep("CORYLUS", da$species)
da$wood[sel] <- 1
sel<-grep("PINUS", da$species)
da$wood[sel] <- 1
sel<-grep("PINE", da$species)
da$wood[sel] <- 1
sel<-grep("POPULUS", da$species)
da$wood[sel] <- 1
sel<-grep("QUERCUS", da$species)
da$wood[sel] <- 1
sel<-grep("ULMUS", da$species)
da$wood[sel] <- 1

# Check for outlier masses
# High values are lorge amounts of litter or trees
da[da$mass.above > 1000 & da$live==1 & da$woo==0,c("year","field", "exp", "plot", "species", "mass.above" )]

# There are a couple species with more than 1500 g of mass that
# are very suspcious based on the species
da <- da[da$mass.above <= 1500, ]


# Read in species attribute look up table
sp.df <- read.csv(here("data/Cedar_Creek_Plant_Taxon_List.csv"),header=T)



names(sp.df)[] <- tolower(names(sp.df))

sp.df<-plyr::rename(sp.df, c("ï..species"="species"))
sp.df$species <- toupper(sp.df$species) 
sp.df$functional.group <- toupper(sp.df$functional.group)
sp.df$duration <- toupper(sp.df$duration)
sp.df$lifeform <- toupper(sp.df$lifeform)
sp.df$pathway <- toupper(sp.df$pathway)
sp.df$origin <- toupper(sp.df$origin)
sp.df$family <- toupper(sp.df$family)

da <- merge(da,sp.df, by='species', all.x=TRUE)

# Set Unknowns to missing
da$origin[da$origin == 'UNKNOWN'] <- NA
da$origin[da$origin == 'NATIVE AND/OR INTRODUCED'] <- NA


da$functional.group[da$functional.group == 'UNKNOW'] <- NA

da$duration[da$duration == 'UNKNOWN'] <- NA
da$duration[da$duration == 'BIENNIAL, PERENNIAL'] <- "BIENNIAL"

# Lump biennials in with annuals as they are pretty similar
# mostly weedy forbs 
unique(da$species[da$duration=="BIENNIAL"])
unique(da$species[da$duration=="BIENNIAL, PERENNIAL"])
unique(da$species[grep('BIENNIAL', da$duration)]) 

da$duration[grep('BIENNIAL', da$duration)] <- "ANNUAL"
unique(da$species[da$duration=="ANNUAL" & da$functional.group=="C4"])
unique(da$species[da$duration=="ANNUAL" & da$functional.group=="F"])

# Add in some species atrributes
sel <- da$species == "QUERCUS RUBRA"
da$functional.group[sel] <- "W"
da$lifeform[sel] <- "WOODY"
da$origin[sel] <- "NATIVE"

# Add in some species atrributes
sel <- da$species == "RUDBEKIA SEROTINA"
da$functional.group[sel] <- "F"
da$lifeform[sel] <- "FORB"
da$origin[sel] <- "NATIVE"

# Add in some species atrributes
sel <- da$species == "POA PRATENSIS"
da$origin[sel] <- "INTRODUCED"

sel <- da$species == "MISCELLANEOUS CAREX SP."
da$functional.group[sel] <- "S"
da$lifeform[sel] <- "SEDGE"
da$origin[sel] <- "NATIVE"


# Capitalize species to get rid of capilization differences in spelling
da$species <- toupper(as.character(da$species))

##deal with some entries where sp. were weighed twice##
summary(freq <- ddply(da[da$live==1 & da$sorted==1, ], .(year, field, exp, plot, subplot, disk, ntrt, nadd, species), colwise(length, .(mass.above)))) ##takes a while##

freq$freq <- freq$mass.above

freq$mass.above <- NULL
freq[freq$freq > 1,]

doubles.df <- merge(freq[freq$freq > 1,], da[c("field", "exp","plot", "year", "species", "mass.above")], by=c("field", "exp","plot", "year", "species"))
doubles.df[c("field", "exp","plot", "year", "species", "mass.above")]

da.mn <- ddply(da, .(field, exp, plot, subplot, year, disk, ntrt, nadd, species, live, sorted, wood, functional.group, lifeform, duration, origin), colwise(mean, .(mass.above)))
##takes a while##

# subset to live, sorted, herbaceous plants
d2 <- da.mn[da.mn$sorted ==1 & da.mn$live ==1 & da.mn$wood==0, c("field", "exp","plot", "subplot", "year", "disk", "ntrt", "nadd", "species", "mass.above")]


#Add some columns####
d3<-d2 %>% mutate(expntrtfieldyear= paste(exp,field, ntrt, year, sep = '_'), expntrtyear= paste(exp,ntrt, year, sep = '_'), ntrt2=ntrt) 

##This data frame (d3) has fields A, B, C, experiment 1 (intact) & exp 2 (disked in 1982) in a long data dataformat ##9 levels of nutrient addition#### years 1982 - 2019 (not every field / exp sampeled in every year## 


# Transpose data to be a site by species matrix
da.wide <- reshape(d3,
                   v.names="mass.above",
                   idvar=c("field", "exp","plot", "subplot", "year", "disk", "ntrt", "nadd", "expntrtyear"),
                   timevar="species",
                   direction="wide") 
# Fill in NA's with zeros in wide data set
da.wide[is.na(da.wide)] <- 0

##This data frame has fields A, B, C, experiment 1 (intact) & exp 2 (disked in 1982) ## years 1982 - 2019 for E002 and years 1982 - 2004 for exp E001 (not every field / exp sampeled in every year##
###wide version format###

# N Treatments (Ntrt) Details (from Table 1 Tilman 1987)
# Ntrt Trt g N/m2/yr Other nutrients 
# 1     A       0.0      All 
# 2     B       1.0      All 
# 3     C       2.0      All 
# 4     D       3.4      All 
# 5     E       5.4      All 
# 6     F       9.5      All 
# 7     G      17.0      All 
# 8     H      27.2      All 
# 9     I       0.0      None 

##make nutrient and years factors#

da.wide$ntrt<-as.factor(da.wide$ntrt)
da.wide$year<-as.factor(da.wide$year)

########Combine east / west subplots into "whole" to match up with rest of data#######

##make only the biomasses the numeric variables##
da.wide$exp<-as.factor(da.wide$exp)
da.wide$disk<-as.factor(da.wide$disk)
da.wide$nadd<-as.factor(as.character(da.wide$nadd))


#subset out the east/ west subplots plots and the whole plots ##

da.wideeastwest<-subset(da.wide, subplot=="East"|subplot=="West", row.names=NULL)
da.widewhole<-subset(da.wide, subplot=="Whole", row.names=NULL)

##Add the values from the subplots together##

EWaddall<-as.data.frame(da.wideeastwest%>%group_by(field,exp,plot,year,disk,ntrt,nadd,expntrtyear, expntrtfieldyear, ntrt2) %>% summarise_if(is.numeric,sum)%>% mutate(subplot="Whole"))


###merge the whole plots back together###
da.wide_allwhole<-rbind(da.widewhole,EWaddall)



dim(da.wide_allwhole) ## dimensions of the dataset = 8798 X 222##

##reorder columns so that all the rows are in the same order across fields etc. 
da.wide5<-da.wide_allwhole%>%arrange(year)%>%arrange(plot)%>%arrange(exp)%>%arrange(field) 


##### Check that the max number of rep within a treatment is 6, since 6 plots in each field and exp recieved the same N treatment##

max(ave(da.wide5$plot, da.wide5$expntrtyear,FUN = seq_along))





##subset to years before 2005## (bc of change in fire regime)

da.wide5$year<-as.numeric(as.character(da.wide5$year)) 

exp12subset<-subset(da.wide5,year<2005)


##subset to selected nutrient treatments below 9.5 g N##

exp12subset_1<-subset(exp12subset,ntrt==9|ntrt==1|ntrt==2|ntrt==4|ntrt==6)

### this data set (exp12subset_1 has fields ABC, E001 and E002 from 1982-2004) ## > 2004 was a change in the fire regime## ONLY the control and selected nutrient treatments below 9.5 g N) ##

###some notes about missing data##
## E001 data set is complete from 1982 - 2004. E002 dataset is missing years 1995, 1998, 2001, 2003 ALL fields##



```



###PERMANOVA AND PCOA###

```{r - Conduct a PCoA using all the veg data across years (log)}


##make nutrient and years factors#

exp12subset_1$ntrt<-as.factor(exp12subset_1$ntrt)
exp12subset_1$year<-as.factor(exp12subset_1$year)


##arrange the dataframe so that that columns and rows are in right order##
exp12subset_1_sorted<-exp12subset_1%>%arrange(year)%>%arrange(plot)%>%arrange(field)%>%arrange(exp)


##just the plot data##
plot<-exp12subset_1_sorted%>% dplyr::select('field':'ntrt2')

##just the veg data##

veg<-exp12subset_1_sorted%>% dplyr::select(`mass.above.ACHILLEA MILLEFOLIUM(LANULOSA)`:`mass.above.VIOLA SP.`)



###log transformation####

vegmatrix<-as.matrix(veg)

logveg<-log(1+vegmatrix)


## Bray curtis dissimiliarty matrix##
distE1andE2<-vegdist(logveg, method="bray")
BCdistall<-as.matrix(distE1andE2)  ##into a matrix#
BCdistalldf<-as.data.frame(BCdistall)  #into a dataframe#

BCdistalldf2<-cbind(plot,BCdistalldf) #merge plot data w/ dissim matrix##

##BCdistalldf2 has the plot info and the BC dissimilarity matrix for the entire experiment (E001 and E002, all three fields combined, with a subsetted selection of the ntrt treatments)

###PcOA for both E011 and E002 together####

PCOA_allfieldE12 <- pcoa(distE1andE2)  ## takes a while##

#PCOA_allfieldE12 <- cmdscale(distE1andE2)  ## Another way of doing the pcoa that gives same results#


## extracting the axis scores for each exp##

PCOAaxesboth <- PCOA_allfieldE12$vectors[,c(1,2)]


##merge the axis scores with the plot data ###

PCOAandplotsboth<-cbind(plot,PCOAaxesboth )

##PCOAandplotsboth has the plot info and the PCOA scores for the entire experiment (E001 and E002, all three fields combined, with a subsetted selection of the ntrt treatments)

```




```{r - Conduct permanova: nutrient, disturbance, field effect (log data)}



####for this analysis, getting rid of the years where not every field was sampeled##

exp12subset_2<-exp12subset_1%>%filter(year!=1995)%>%filter(year!=1998)%>%filter(year!=2001)%>%filter(year!=2003)



###log transform the biomass data###

justveg<-as.matrix(exp12subset_2[,12:222])

logveg<-log(1+justveg)

exp12subset_2<-cbind(exp12subset_2[,1:11], logveg)


#drop the levels not sampeled

levels(exp12subset_2$year)
exp12subset_2$year <- as.factor(as.character(exp12subset_2$year))

### Extract one year to get the apply function to work#####

da.wide_1990<-subset(exp12subset_2, year=="1990", row.names=NULL)

##Do a PERMANOVA looping over years, looking at effect of exp (disturbed or not), ntrt treatment# and field (A, B, C)###

###
permallfields<-lapply(X = split.data.frame(x = exp12subset_2, f=exp12subset_2$year),function(t){
  z<-adonis2(vegdist(t[,11:175])~exp+ntrt+field, data= da.wide_1990)
  return(data.frame(z))
  }
  )



##dataframe of results##
permmaster<-plyr::ldply(permallfields, data.frame)
names<-c("Disturbance", "Fertilization", "Field", "residual", "total")

sourcevariation<-rep(names,length(unique(permmaster$.id)))

permmaster2<-cbind(sourcevariation,permmaster)
  
permmaster3<-subset(permmaster2,sourcevariation=="Disturbance"|sourcevariation=="Fertilization"|sourcevariation=="Field") 

##This dataframe (permmmaster3) has the results from a permanova for each independent variable (disturbance, field, fertilization) for every year in the dataset w/ everything sampeled. The R2 column is what is plotted in Fig 1 ###



```



```{r - recreate FIG 1, R2 over time dist, nut, field (log data)}


permmaster4<- permmaster3 %>% mutate (year=as.numeric(.id))  %>% mutate (year2=year-1981)

permmaster4$sourcevariation<- factor(permmaster4$sourcevariation, levels = c("Disturbance", "Fertilization", "Field"))

my_plotlabels = paste0('(', letters, ') ')

abclabels = paste0(my_plotlabels[1:length(levels(permmaster4$sourcevariation))]) 

permmaster5<-cbind(permmaster4, abclabels)


##make the plot##

mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))+ theme(panel.grid.major = element_line(colour = "gray 85"))+ theme(panel.grid.minor = element_line(colour = "gray 95"))




variancetimeall<-permmaster5 %>% ggplot(aes(x=as.numeric(year2), y=R2, group=sourcevariation)) +
    geom_point() +
    facet_wrap(~sourcevariation, strip.position = c("top"))+
    geom_smooth(method="loess", colour="black", size=0.5, fill = "gray1") +
    geom_text(aes(x = 2.0, y = 0.49, label = abclabels, group =     abclabels),size = 4,check_overlap = T)+
    ylab(expression('Explained variation'))+
    xlab("Time since experiment (years)")+
    mytheme+
    theme(strip.text = element_text(face="bold", size=12))+
     theme(strip.background = element_blank())

   ###

```



```{r - recreate FIG 2, PCOA ordination by year (log data)}



mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=8, colour = "black"))+theme(axis.text.y=element_text(size=8, colour = "black"))+theme(axis.title=element_text(size=10))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))



##PCOAandplotsE001 (and E002) have 11 colummns of plot data, PC0A axis 1 and 2 for each plot (in each year, field, subset of ntrt conditions)##


##all fields. averaging across plots in all fields (n=18) with the same treatment group##

#first check that there are 18 in each ntrt treatment / year##
PCOAandplotscount<-PCOAandplotsboth %>% count(year,disk,ntrt)

####2003 for E002 only has 1 field so we remove it###

PCOAandplotsboth2<-PCOAandplotsboth %>%filter(year!=2003)

PCOAandplotscount2<-PCOAandplotsboth2 %>% count(year,disk,ntrt)##good now###

PCOAandplotsboth2$Axis.1<-as.numeric(PCOAandplotsboth2$Axis.1)
PCOAandplotsboth2$Axis.2<-as.numeric(PCOAandplotsboth2$Axis.2)

PCOAandplotsboth2$year<-as.numeric(as.character(PCOAandplotsboth2$year))



PCOAandplotsavg<-PCOAandplotsboth2 %>% group_by(exp,year,disk,ntrt) %>% dplyr::summarise(meanAxis1=mean(Axis.1, na.rm=TRUE),sdAxis1=sd(Axis.1, na.rm=TRUE),seAxis1=sd(Axis.1, na.rm=TRUE)/sqrt(n()),meanAxis2=mean(Axis.2, na.rm=TRUE),sdAxis2=sd(Axis.2, na.rm=TRUE),seAxis2=sd(Axis.2, na.rm=TRUE)/sqrt(n()), n=n())



####

## plot specifications

PCOAandplotsavg2<-mutate(PCOAandplotsavg, ntrt2=ntrt)

PCOAandplotsavg2$ntrt2<-PCOAandplotsavg2$ntrt

PCOAandplotsavg2$ntrt2  <- plyr::mapvalues(PCOAandplotsavg2$ntrt2, from=c("9", "1", "2", "3", "4","5","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))

PCOAandplotsavg2$ntrt2 <- factor(PCOAandplotsavg2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))

PCOAandplotsavg2<-PCOAandplotsavg2  %>% mutate(yearlabel=ifelse(year==1982|year==1992|year==2004,year, NA))%>%mutate(exp_ntrt= as.factor(paste(exp, ntrt, sep = '_')))

PCOAandplotsavg2$exp_ntrt <- factor(PCOAandplotsavg2$exp_ntrt , levels = c("1_9", "2_9", "1_1","2_1", "1_2", "2_2", "1_4", "2_4", "1_6", "2_6"))

PCOAandplotsavg3<-PCOAandplotsavg2 %>% mutate(abclabel=exp_ntrt)
levels(PCOAandplotsavg3$abclabel) <- c( "(a) ", "(b) ", "(c) ", "(d) ", "(e) ", "(f) ", "(g) ", "(h) ", "(i) ", "(j) ")

## plot specifications
PCOAandplotsavg3$ntrt2  <- plyr::mapvalues(PCOAandplotsavg3$ntrt2, from=c("9", "1", "2", "3", "4","5","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))
PCOAandplotsavg3$ntrt2 <- factor(PCOAandplotsavg3$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))


PCOAandplotsavg3$disk<-plyr::mapvalues(PCOAandplotsavg3$disk, from=c("0", "1"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))




##
##to draw points between years###

ann_textdisk3 <- data.frame(meanAxis1 = c(0.00, 0.00), meanAxis2 =c(0.27, 0.27),exp= c("Intact in 1982 (E001)","Disturbed in 1982 (E002)"), year=c(1982, 1982), exp_ntrt=c("1_9", "2_9"),abclabel = factor(c("(a) ","(b) " ),levels = c("(a) ", "(b) ", "(c) ", "(d) ", "(e) ", "(f) ", "(g) ", "(h) ", "(i) ", "(j) ")))

ann_textntrt <- data.frame(meanAxis1 = c(0.45, 0.45, 0.45, 0.45, 0.45), meanAxis2 =c(0.05, 0.05, 0.05, 0.05, 0.05),exp= c("2","2","2","2","2"), year=c(1982, 1982, 1982, 1982, 1982), exp_ntrt=c("2_9", "2_1", "2_2", "2_4", "2_6"),abclabel = factor(c("(b) ","(d) ", "(f) ","(h) ","(j) "),levels = c("(a) ", "(b) ", "(c) ", "(d) ", "(e) ", "(f) ", "(g) ", "(h) ", "(i) ", "(j) ")))



PcoAfieldAveragedE001E002<-ggplot(PCOAandplotsavg3[order(PCOAandplotsavg3$year),], aes(x=meanAxis1  , y=meanAxis2 , group=exp_ntrt, color=year, label=year)) +
    geom_point(size=0.75) +
    geom_errorbar(aes(ymin=meanAxis2-seAxis2, ymax=meanAxis2+seAxis2), color="dark gray", size=0.25) +
    geom_errorbarh(aes(xmin=meanAxis1-seAxis1, xmax=meanAxis1+seAxis1), color="dark gray", size=0.25) +
    geom_path(aes(x=meanAxis1 , y=meanAxis2 , group=exp_ntrt, color=year),size=0.5)+
    geom_point(size=0.75) +
    xlim(c(min(PCOAandplotsavg3$meanAxis1-PCOAandplotsavg3$seAxis1)),max(PCOAandplotsavg3$meanAxis1+PCOAandplotsavg3$seAxis1))+
    ylim(c(min(PCOAandplotsavg3$meanAxis2-PCOAandplotsavg3$seAxis2)),max(PCOAandplotsavg3$meanAxis2+PCOAandplotsavg3$seAxis2))+
    ylab("PCoA 2")+
    xlab("PCoA 1")+
    facet_grid(ntrt2~disk) + 
    #geom_text(aes(label=yearlabel),position=position_jitter(width=0.05,height=0.05), color="black", size=2.5)+
    mytheme+
    scale_color_viridis(option = "D")+
    scale_fill_viridis(option = "D",trans = 'reverse')+
    guides(#reverse color order (higher value on top)
    color = guide_colorbar(reverse = TRUE),
    #reverse size order (higher diameter on top) 
    size = guide_legend(reverse = TRUE))+
   #scale_y_continuous(breaks=c(-0.2, 0.0, 0.2, 0.4))+
    geom_text(aes(x = -0.26, y = 0.16, label = abclabel, group =     abclabel),size = 3,check_overlap = T)+
    theme(strip.background = element_blank())+
    theme(strip.text = element_text(face="bold", size=10))+
    theme(panel.spacing.y=unit(0.5, "lines"))
    #geom_text_repel(aes(label=yearlabel),size = 2.5, color="black")


```



```{r - recreate FIG 4 cd, PCoA by decade (log data)}


PCOAandplotsavg<-PCOAandplotsboth %>% group_by(exp,year,disk,ntrt) %>% dplyr::summarise(meanAxis1=mean(Axis.1, na.rm=TRUE),sdAxis1=sd(Axis.1, na.rm=TRUE),seAxis1=sd(Axis.1, na.rm=TRUE)/sqrt(n()),meanAxis2=mean(Axis.2, na.rm=TRUE),sdAxis2=sd(Axis.2, na.rm=TRUE),seAxis2=sd(Axis.2, na.rm=TRUE)/sqrt(n()), n=n())


###get rid of years where we don't have all 18 plots###
####2003 for E002 only has 1 field, so remove it###

PCOAandplotsavg2<-subset(PCOAandplotsavg, n==18)

PCOAandplotsavg2<-PCOAandplotsavg2%>%mutate(ntrt2 = as.factor(ntrt))

PCOAandplotsavg2<-PCOAandplotsavg2  %>% mutate(yearlabel=ifelse(year==1982|year==1988|year==1992|year==2004,year, NA))

######

##by decade####

PcoAplaying <- PCOAandplotsavg2 %>% filter(year %in% c(1982,2004))

arrow_start <- PCOAandplotsavg2 %>% filter(year ==1982)
arrow_stop<- PCOAandplotsavg2 %>% filter(year == 2004)

wide_PcoAplaying <- pivot_wider(PcoAplaying, names_from = year, values_from = year)

##just pull out the first year of data##

PCOAandplots1982<-subset(PCOAandplotsavg2, year==1982)

##pull out the first, last, and midpoint year##

PCOAandplots198219922004<-subset(PCOAandplotsavg2, year==1982|year==1992|year==2004)

PCOAandplots198219922004<-PCOAandplots198219922004 %>% mutate(abclabel=exp)%>% mutate(ntrt2=ntrt) %>%  mutate(expntrt= paste(exp, ntrt, sep = '_'))

levels(PCOAandplots198219922004$abclabel) <- c( "(a) ", "(b) ")


PCOAandplots198219922004$ntrt2  <- mapvalues(PCOAandplots198219922004$ntrt2, from=c("9", "1", "2", "3", "4","5","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))
PCOAandplots198219922004$ntrt2 <- factor(PCOAandplots198219922004$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))


#first arrow and second##
PcoAplayingfirst <- PCOAandplotsavg2 %>% filter(year %in% c(1982,1992, 2004))

PcoAplayingfirst2<-PcoAplayingfirst %>% mutate(abclabel=exp)
levels(PcoAplayingfirst2$abclabel) <- c( "(c) ", "(d) ")

#make sure its in the right order##
PcoAplayingfirst2<-PcoAplayingfirst2 %>% arrange(year)%>% arrange(exp)%>% arrange(ntrt)%>% mutate(ntrt2=ntrt) %>%  mutate(expntrt= paste(exp, ntrt, sep = '_'))



PcoAplayingfirst2$ntrt2  <- mapvalues(PcoAplayingfirst2$ntrt2, from=c("9", "1", "2", "3", "4","5","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))
PcoAplayingfirst$ntrt2 <- factor(PcoAplayingfirst$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))

wide_PcoAplaying1 <- pivot_wider(PcoAplayingfirst2, names_from = year, values_from = year)

wide_PcoAplaying1$`1982`<-as.numeric(as.character(wide_PcoAplaying1$`1982`))
wide_PcoAplaying1$`1992`<-as.numeric(as.character(wide_PcoAplaying1$`1992`))
wide_PcoAplaying1$`2004`<-as.numeric(as.character(wide_PcoAplaying1$`2004`))


PCOAandplotslabels<-subset(PcoAplayingfirst2, ntrt==4)
PCOAandplotslabels<-PCOAandplotslabels%>%  mutate(expntrt= paste(exp, ntrt, sep = '_'))%>%  mutate(yearlabel= year)


######

## plot specifications
PCOAandplotsavg2$ntrt2  <- mapvalues(PCOAandplotsavg2$ntrt2, from=c("9", "1", "2", "3", "4","5","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))
PCOAandplotsavg2$ntrt2 <- factor(PCOAandplotsavg2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))


PCOAandplotsavg2$disk<-mapvalues(PCOAandplotsavg2$disk, from=c("0", "1"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))



ann_textdisk2 <- data.frame(meanAxis1 = c(0.03, 0.03), meanAxis2 =c(0.27,0.27),exp= c("Intact in 1982 (E001)","Disturbed in 1982 (E002)"), ntrt2=c("None", "None"), expntrt=c("1_9" ,"2_9"), abclabel = factor(c("(a) ","(b) " ),levels = c("(c) ", "(d) ")))


####
mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))+ theme(panel.grid.major = element_line(colour = "gray 85"))+ theme(panel.grid.minor = element_line(colour = "gray 95"))


PcOAdecade<-ggplot(data=PcoAplayingfirst2, aes(x=meanAxis1  , y=meanAxis2 , color=ntrt2)) +
  geom_hline(yintercept = 0)+geom_vline(xintercept = 0)+
  geom_point(data=PCOAandplots198219922004, aes(x=meanAxis1  , y=meanAxis2 , group=ntrt2, shape=as.factor(year)), size=2)+
    geom_errorbar(data=PCOAandplots198219922004, aes(ymin=meanAxis2-seAxis2, ymax=meanAxis2+seAxis2,color=ntrt2), size=0.25) +
  geom_errorbarh(data=PCOAandplots198219922004, aes(xmin=meanAxis1-seAxis1, xmax=meanAxis1+seAxis1,color=ntrt2),  size=0.25) +
  geom_point(data=PCOAandplots198219922004, aes(x=meanAxis1  , y=meanAxis2 , group=ntrt2,color=ntrt2,shape=as.factor(year)))+
  geom_segment(data=wide_PcoAplaying1, aes(x=`1982`  , y=`1982`, xend =`1992`, yend = `1992`, color=ntrt2), arrow=arrow(length=unit(0.0, "cm"))) +
    geom_path(size=0.75, arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))+
      geom_segment(data=wide_PcoAplaying1, aes(x=`1992`  , y=`1992`, xend =`2004`, yend = `2004`, color=ntrt2), arrow=arrow(length=unit(0.0, "cm"))) +
    geom_path(size=0.75, arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))+
          geom_text(data=PCOAandplotslabels, aes(label=yearlabel),position = position_nudge(y = -0.018, x=-0.02), color="black", size=3.0)+
    ylab("PCoA 2")+
    xlab("PCoA 1")+
    facet_grid(~disk) + 
    xlim(-0.28,0.375)+
    ylim(-0.35,0.15)+
    scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
    mytheme+
    theme(legend.position="none")+
    scale_shape_manual(values = c(21, 22, 24))+
      theme(strip.background = element_blank())+
      geom_text(aes(x = -0.25, y = 0.14, label = abclabel, group =     abclabel),size = 4, color= "black", check_overlap = T)+theme(strip.text = element_blank())


```



###DISPERSION####
```{r - Calculate the distance TO centroid within ntrt treatments (log data)}


###need to have the most recent verion of vegan from github for this to work##
#https://github.com/vegandevs/vegan/issues/331#


#######################

####for this analysis, getting rid of the years where not every field was sampeled##

exp12subset_2<-exp12subset_1%>%filter(year!=1995)%>%filter(year!=1998)%>%filter(year!=2001)%>%filter(year!=2003)

##just the veg data##

veg<-exp12subset_2%>% dplyr::select(`mass.above.ACHILLEA MILLEFOLIUM(LANULOSA)`:`mass.above.VIOLA SP.`)



###log transformation####

vegmatrix<-as.matrix(veg)

logveg<-log(1+vegmatrix)



##using the apply function to calculation the dispersion (beta disp) for each nutrient treatment for each year#

bd_allboth<-betadisper(vegdist(logveg),exp12subset_2$expntrtyear, type = "centroid")    ####Note that you have to specify the centroid ##  ##takes a bit##

#bd_allbothEUC<-betadisper(vegdist(logveg, method="euclidean"),exp12subset_2$expntrtyear, type = "centroid")    ##same thing but in Euclidean distance###

##some squared distances are negative and changed to zero##


######making group distances into a dataframe with rownames###

bd_allboth_2<-as.data.frame(bd_allboth$group.distances, row.names = rownames(bd_allboth$group.distances))

setDT(bd_allboth_2, keep.rownames = TRUE)[]


bd_allboth_df<-separate(data = bd_allboth_2, col = rn, into = c("exp", "ntrt", "year"))

bd_allboth_df$year<-as.numeric(bd_allboth_df$year)
bd_allboth_df$ntrt<-as.factor(bd_allboth_df$ntrt)

bd_allboth_df$distances<-as.numeric(bd_allboth_df$`bd_allboth$group.distances`)

bd_allboth_df2<-bd_allboth_df%>%mutate(ntrt2 = ntrt)

###this dataframe (bd_allboth_df2) has the average distance to the centroid for each disurbance_nutrient group for each year. Distances = the average Bray dist for 18 plots to their group centroids (what is plotted in Fig 3 a) ##


```


```{r - AIC table dist TO centroid, TABLE S1 }

####test whether the distance to centroid is best fit by a linear or saturating function##


head(bd_allboth_df2)

bd_allboth_df2<-bd_allboth_df2%>% mutate (year2=year-1981)


###linear fit###

distlin<-lm(distances~year2, data=bd_allboth_df2)


##saturating function##
distnonlin<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = bd_allboth_df2)


rawaic<-AIC(distlin, distnonlin)
nR<-dim(bd_allboth_df2)[1]  #Sample size 
aictable(rawaic,nR)

###nonlinear is the best fit###

```


###check this one####
```{r - model fits and error dist TO centroid, TABLE S2}


head(bd_allboth_df2)

###https://www.statforbiology.com/nonlinearregression/usefulequations#asymptotic_regression_model###

groupdistancesboth2_df<-bd_allboth_df2%>% mutate(year2=year-1981)

groupdistcontrol<-bd_allboth_df2%>% filter(ntrt==9)


###E002 seperate by nutrient and experiment####

controlE002<-subset(groupdistancesboth2_df, ntrt==9 & exp==2)
E002distntrt1<-subset(groupdistancesboth2_df, ntrt==1 & exp==2)
E002distntrt2<-subset(groupdistancesboth2_df, ntrt==2 & exp==2)
E002distntrt4<-subset(groupdistancesboth2_df, ntrt==4 & exp==2)
E002distntrt6<-subset(groupdistancesboth2_df, ntrt==6 & exp==2)

#
#E002 #Non linear model fits##

controlE002fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = controlE002)
E002distntrt1fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E002distntrt1)
E002distntrt2fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E002distntrt2)
E002distntrt4fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E002distntrt4)
E002distntrt6fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E002distntrt6)


##Seperate experiment and N treatments, E001 ##

controlE001<-subset(groupdistancesboth2_df, ntrt==9 & exp==1)
E001distntrt1<-subset(groupdistancesboth2_df, ntrt==1 & exp==1)
E001distntrt2<-subset(groupdistancesboth2_df, ntrt==2 & exp==1)
E001distntrt4<-subset(groupdistancesboth2_df, ntrt==4 & exp==1)
E001distntrt6<-subset(groupdistancesboth2_df, ntrt==6 & exp==1)


####Nonlinear model fits E001##

controlE001fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = controlE001)
E001distntrt1fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E001distntrt1)
E001distntrt2fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E001distntrt2)
E001distntrt4fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E001distntrt4)
#E001distntrt6fit<- nls(distances~ SSasymp(year2, Asym, R0, lrc), data = E001distntrt6) ##doesn't work...see below###

####this (ntrt6) is weird (have to specify lrc and then it works...##

x6=E001distntrt6$year2
y6=E001distntrt6$distances

##first, specify the fit##
asympfunc <- function(x6, Asym, R0, lrc=coef(E002distntrt6fit)[3])
    Asym + (R0 - Asym) * exp(-exp(lrc) * x6)
nls(y6 ~ asympfunc(x6, Asym, R0),
    data = data.frame(x6, y6),
    start = list(Asym = 0.3,R0=0.5))


E001distntrt6fit<-nls(y ~ asympfunc(x6, Asym, R0),
    data = data.frame(x6, y6),
    start = as.list(coef(E002distntrt6fit)[c(1, 2)]))

###THESE ARE ALL THE VALUES IN TABLE S1###

summary(controlE001fit)
summary(E001distntrt1fit)
summary(E001distntrt2fit)
summary(E001distntrt4fit)
summary(E001distntrt6fit)

summary(controlE002fit)
summary(E002distntrt1fit)
summary(E002distntrt2fit)
summary(E002distntrt4fit)
summary(E002distntrt6fit)





```


```{r - recreate FIG 3ab, Distance to the centroid (log data)}


head(groupdistancesboth2_df)


##confidence intervals around the control only###

## confidence intervals E001##

confE001 <- as.data.frame(predFit(controlE001fit,interval = "confidence", level= 0.95))

confE001_1<-cbind(controlE001,confE001)

## confidence intervals E002##

confE002 <- as.data.frame(predFit(controlE002fit,interval = "confidence", level= 0.95))

confE002_2<-cbind(controlE002,confE002)

confboth<-rbind(confE001_1,confE002_2)

##combine and add label###
confboth<-rbind(confE001_1, confE002_2)

confboth_2<-confboth %>% mutate(abclabel=exp)
levels(confboth_2$abclabel) <- c( "(a) ", "(b) ")



##### plot specifications###

groupdistcontrol<-groupdistcontrol %>% mutate(abclabel=exp)
levels(groupdistcontrol$abclabel) <- c("(a) ", "(b) ")


groupdistancesboth2_df<-groupdistancesboth2_df %>% mutate(abclabel=exp)
levels(groupdistancesboth2_df$abclabel)<- c("(a) ", "(b) ")


confboth_2$exp<-mapvalues(confboth_2$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

confboth_2$exp<- factor(confboth_2$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


groupdistcontrol$exp<-mapvalues(groupdistcontrol$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

groupdistcontrol$exp<- factor(groupdistcontrol$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))




groupdistancesboth2_df$exp<-mapvalues(groupdistancesboth2_df$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

groupdistancesboth2_df$exp<- factor(groupdistancesboth2_df$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

groupdistancesboth2_df<-groupdistancesboth2_df%>%mutate(ntrt2 = as.factor(ntrt))

groupdistancesboth2_df$ntrt2  <- mapvalues(groupdistancesboth2_df$ntrt2, from=c("9", "1", "2", "3", "4","5","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))
groupdistancesboth2_df$ntrt2 <- factor(groupdistancesboth2_df$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))


groupdistancesboth2_df<-groupdistancesboth2_df %>% mutate(abclabel=exp)
levels(groupdistancesboth2_df$abclabel) <- c( "(a) ", "(b) ")


ann_textdisk1 <- data.frame(year2 = c(12,12), distances =c(0.7, 0.7),exp= c("Intact in 1982 (E001)","Disturbed in 1982 (E002)"), ntrt2=c("None", "None"), abclabel = factor(c("(a) ","(b) " ),levels = c("(a) ", "(b) ")))

####



mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title.x=element_text(size=12))+theme(axis.title.y=element_text(size=10))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))+ theme(panel.grid.major = element_line(colour = "gray 85"))+ theme(panel.grid.minor = element_line(colour = "gray 95"))


disttocentroidplot<- ggplot(data=groupdistancesboth2_df) +
    geom_point(data=groupdistancesboth2_df, aes(x=as.numeric(as.character(year2)), y=distances, group=ntrt2, color=ntrt2)) +
    geom_point(data=groupdistcontrol, aes(x=as.numeric(as.character(year2)), y=distances),     color="gray45") +
  geom_smooth(data=groupdistancesboth2_df,aes(x=as.numeric(as.character(year2)), y=distances, color=ntrt2), method="nls",formula=y ~ a * x^b,method.args=list(start=c(a=0.5,b=-0.1)),se=FALSE, size = 0.5)+ 
  geom_smooth(data=groupdistcontrol,aes(x=as.numeric(as.character(year2)), y=distances), method="nls",formula=y ~ a * x^b,method.args=list(start=c(a=0.5,b=-0.1)),se=FALSE,colour = "gray45", size = 0.5)+
  geom_ribbon(data=confboth_2, aes(ymin=lwr, ymax=upr, x=year2), color=NA, alpha=0.2)+
    facet_grid(~exp)+ 
    ylab(expression('Average distance to group centroid'))+
    xlab("Time since experiment")+
    mytheme+
    theme(legend.position="right")+
    theme(legend.title = element_blank()) + 
    scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
    theme(strip.text = element_text(size=12, face="bold"))+
     theme(strip.background = element_blank())+
     geom_text(aes(x = 1.5, y = 0.65, label = abclabel, group =     abclabel),size = 4,check_overlap = T)

```


```{r - Calculate the distance BTWN centroids (logdata) }

##pulling out the group centroids by year####

centroiddf<-as.data.frame(bd_allboth[["centroids"]])

centroiddf_2<-centroiddf[,1:2]

centroiddf_2 <- tibble::rownames_to_column(centroiddf_2, "expntrtyear")

centroiddf_2<-separate(data = centroiddf_2, col = expntrtyear, into = c("exp", "ntrt", "year"))

centroiddf_3<-centroiddf_2%>% mutate(expyear= paste(exp,year, sep = '_')) 


###grand mean centroid for all treatments each year###

grandcentroid<-centroiddf_2%>% group_by(exp, year) %>% dplyr::summarise(PcoA1centroid=mean(PCoA1, na.rm=TRUE),PcoA2centroid=mean(PCoA2, na.rm=TRUE))%>% mutate(expyear= paste(exp,year, sep = '_')) 


### merge the treatments means with the grand mean centroid for each year##

centroiddf_3<-merge(data.frame(centroiddf_3), data.frame(grandcentroid), by = "expyear", all = TRUE)

###calculate the distance from each centroid to grand centroid using pythagoreum theorem###

centroiddf_4<-centroiddf_3 %>% mutate(distgrandcent=sqrt((PcoA1centroid- PCoA1)^2+(PcoA2centroid-PCoA2)^2))


###average distance of each group to its grand mean##

centroiddists<-centroiddf_4%>% group_by(expyear) %>% dplyr::summarise(meandist=mean(distgrandcent, na.rm=TRUE))%>%separate(col = expyear, into = c("exp", "year"))


##this dataframe (centroidsdists) has the average distance between nutrient group centroids for each year, for each experiment####

```


```{r - AIC table dist BTWN centroid, TABLE S3 }


centroiddists2<-centroiddists%>% mutate (year2=as.numeric(as.character(year))-1981)

###linear fit###

distbtwnlin<-lm(meandist~year2, data=centroiddists2)


##saturating function##
distbtwnnonlin<- nls(meandist~ SSasymp(year2, Asym, R0, lrc), data = centroiddists2)


rawaic<-AIC(distbtwnlin, distbtwnnonlin)
nR<-dim(bd_allboth_df2)[1]  #Sample size 
aictable(rawaic,nR)

###Nonlinear is the best fit###

```


```{r - recreate FIG 3 cd, distance BTWN centroids (logdata)}

centroiddists$year<-as.numeric(as.character(centroiddists$year))

centroiddists_2<-centroiddists%>% mutate(year2=year-1981)


####subset by exp###

centroiddists_E001<-subset(centroiddists_2, exp==1)
centroiddists_E002<-subset(centroiddists_2, exp==2)


###https://www.r-bloggers.com/2021/07/asymptotic-confidence-intervals-for-nls-regression-in-r/##

asympfitcentroidE001<- nls(meandist~ SSasymp(year2, Asym, R0, lrc), data = centroiddists_E001)

asympfitcentroidE002 <- nls(meandist~ SSasymp(year2, Asym, R0, lrc), data = centroiddists_E002)


###confidence intervals around fit###


new.dataE001 <- data.frame(year2=centroiddists_E001$year2)
new.dataE002 <- data.frame(year2=centroiddists_E002$year2)


confE001_centroid <- as_tibble(predFit(asympfitcentroidE001, newdata = new.dataE001, interval = "confidence", level= 0.95)) %>% mutate(year2 = new.dataE001$year2)

confE001_centroiddf<-as.data.frame(confE001_centroid)

confE001_centroiddf2<-cbind(centroiddists_E001, confE001_centroiddf)


confE002_centroid <- as_tibble(predFit(asympfitcentroidE002, newdata = new.dataE002, interval = "confidence", level= 0.95)) %>% mutate(year2 = new.dataE002$year2)


confE002_centroiddf<-as.data.frame(confE002_centroid)

confE002_centroiddf2<-cbind(centroiddists_E002, confE002_centroiddf)


      ##combine togeter and add abc label##
confcentroidboth<-rbind(confE001_centroiddf2, confE002_centroiddf2)


###for plotting####


centroiddists_2$exp<-mapvalues(centroiddists_2$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

centroiddists_2$exp<- factor(centroiddists_2$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


centroiddists_3<-centroiddists_2%>% mutate(abclabel=exp)

levels(centroiddists_3$abclabel) <- c( "(c) ", "(d) ")


confcentroidboth$exp<-mapvalues(confcentroidboth$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

confcentroidboth$exp<- factor(confcentroidboth$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


confcentroidboth_2<-confcentroidboth[,1:7] %>% mutate(abclabel=exp)

levels(confcentroidboth_2$abclabel) <- c( "(c) ", "(d) ")



distbetweencentroids<-ggplot(data=centroiddists_3,aes(x=as.numeric(as.character(year2)),y=meandist)) +
    geom_point() +
    facet_grid(~exp) +
    ylab(expression('Average distance between group centroids'))+
    xlab("Time since experiment (years)")+
    geom_smooth(method="nls", formula=y~SSasymp(x, Asym, R0, lrc), colour = "black", size = 0.5, se=F, fullrange=T)+
    geom_ribbon(data=confcentroidboth_2, aes(ymin=lwr, ymax=upr), color=NA, fill = "gray1", alpha=0.2)+
    mytheme+
    theme(legend.position="none")+
     theme(strip.text = element_blank())+
     theme(strip.background = element_blank())+
     geom_text(aes(x = 1.5, y = 0.18, label = abclabel, group =     abclabel),size = 4,check_overlap = T)
  



```


```{r - Model fits and error dist BTWN centroid, TABLE S4}


#dist between centroids#
asympfitcentroidE001
asympfitcentroidE002

summary(asympfitcentroidE001)
summary(asympfitcentroidE002)

```


```{r - recreate FIG 3 abcd}

quaddisttocentroidplotnox<- ggplot(data=groupdistancesboth2_df) +
    geom_point(data=groupdistancesboth2_df, aes(x=as.numeric(as.character(year2)), y=distances, group=ntrt2, color=ntrt2)) +
    geom_point(data=groupdistcontrol, aes(x=as.numeric(as.character(year2)), y=distances),     color="gray45") +
  geom_smooth(data=groupdistancesboth2_df,aes(x=as.numeric(as.character(year2)), y=distances, color=ntrt2), method="nls",formula=y ~ a * x^b,method.args=list(start=c(a=0.5,b=-0.1)),se=FALSE, size = 0.5)+ 
  geom_smooth(data=groupdistcontrol,aes(x=as.numeric(as.character(year2)), y=distances), method="nls",formula=y ~ a * x^b,method.args=list(start=c(a=0.5,b=-0.1)),se=FALSE,colour = "gray45", size = 0.5)+
  geom_ribbon(data=confboth_2, aes(ymin=lwr, ymax=upr, x=year2), color=NA, alpha=0.2)+
    facet_grid(~exp)+ 
    ylab(expression('Average distance to group centroid'))+
    xlab("Time since experiment")+
    mytheme+
    theme(legend.position="right")+
    theme(legend.title = element_blank()) + 
    scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
    theme(strip.text = element_text(size=12, face="bold"))+
     theme(strip.background = element_blank())+
     geom_text(aes(x = 1.5, y = 0.65, label = abclabel, group =     abclabel),size = 4,check_overlap = T)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


disttocentroidplot_2<-quaddisttocentroidplotnox+ theme(legend.position = c(0.85, 0.75))+theme(legend.text = element_text(size = 8))+ guides(color = guide_legend(override.aes = list(size = 0.5)))+theme(legend.margin = margin(0, 0, 0, 0))+theme(legend.spacing.y = unit(0, "mm"))+ theme(legend.key.size = unit(1.5, "mm"))+ scale_y_continuous(labels = scales::number_format(accuracy = 0.01))


    
g11 <- ggplotGrob(disttocentroidplot_2)
g22 <- ggplotGrob(distbetweencentroids)

grid.arrange(g11, g22 , ncol = 1, heights = c(1, 1))

```



###SINOUSITY####
```{r - calculate sinuosity of trajectories (log data)}

#https://cran.rstudio.com/web/packages/trajr/vignettes/trajr-vignette.html#


head(PCOAandplotsboth)


##make each experiment, nutrient, plot, its own combo###
PCOAandplotsboth2<-PCOAandplotsboth%>% mutate(expntrtplot= paste(exp, ntrt, plot, sep = '_'))

PCOAandplotsboth3<-PCOAandplotsboth2 %>%  ungroup() %>% dplyr::select(expntrtplot, Axis.1, Axis.2, year, -exp, -disk)


###calculate sinuoisty for each plot using the traj from coords function##
trajectoriesplot<-lapply(X = split.data.frame(x = PCOAandplotsboth3, f=PCOAandplotsboth3$expntrtplot),function(t){
  PCOAscores<-t %>%   ungroup() %>% dplyr::select(Axis.1, Axis.2, year)
  trajs <- TrajFromCoords(PCOAscores)
  sinuosity<-TrajSinuosity2(trajs)
  return(data.frame(sinuosity))
  }
  )

##dataframe of results##
trajectoriesplotdf<-plyr::ldply(trajectoriesplot, data.frame)


trajectoriesplotdf_2<-separate(data = trajectoriesplotdf, col = .id, into = c("exp", "ntrt", "plot"))

##mean sinuoisty for each exp_ntrt###

sinuosityavg<-trajectoriesplotdf_2%>% group_by(exp, ntrt) %>% dplyr::summarise(sinuosityavg=mean(sinuosity, na.rm=TRUE),sdsinuosiy=sd(sinuosity, na.rm=TRUE),sesinuosity=sd(sinuosity, na.rm=TRUE)/sqrt(n()), uprconfint=sinuosityavg+(1.96*sesinuosity),lwrconfint=sinuosityavg-(1.96*sesinuosity), n=n())


```


```{r - calculate sinuosity two decades}


###split by decade##

PCOAandplotsfirstdecade<-subset(PCOAandplotsboth, as.numeric(as.character(year))<=1992)

PCOAandplotsseconddecade<-subset(PCOAandplotsboth, as.numeric(as.character(year))>1992)


##make each experiment, nutrient, plot, its own combo###
PCOAandplotsfirstdecade2<-PCOAandplotsfirstdecade%>% mutate(expntrtplot= paste(exp, ntrt, plot, sep = '_'))

PCOAandplotsfirstdecade3<-PCOAandplotsfirstdecade2 %>%  ungroup() %>% select(expntrtplot, Axis.1, Axis.2, year, -exp, -disk)


PCOAandplotsseconddecade2<-PCOAandplotsseconddecade%>% mutate(expntrtplot= paste(exp, ntrt, plot, sep = '_'))

PCOAandplotsseconddecade3<-PCOAandplotsseconddecade2 %>%  ungroup() %>% select(expntrtplot, Axis.1, Axis.2, year, -exp, -disk)

###calculate sinuoisty for each plot###

###
trajectoriesplotfirst<-lapply(X = split.data.frame(x = PCOAandplotsfirstdecade3, f=PCOAandplotsfirstdecade3$expntrtplot),function(t){
  PCOAscores<-t %>%   ungroup() %>% select(Axis.1, Axis.2, year)
  trajs <- TrajFromCoords(PCOAscores)
  sinuosity<-TrajSinuosity2(trajs)
  return(data.frame(sinuosity))
  }
  )


trajectoriesplotsecond<-lapply(X = split.data.frame(x = PCOAandplotsseconddecade3, f=PCOAandplotsseconddecade3$expntrtplot),function(t){
  PCOAscores<-t %>%   ungroup() %>% select(Axis.1, Axis.2, year)
  trajs <- TrajFromCoords(PCOAscores)
  sinuosity<-TrajSinuosity2(trajs)
  return(data.frame(sinuosity))
  }
  )



##dataframe of results##
trajectoriesplotdffirst<-plyr::ldply(trajectoriesplotfirst, data.frame)
trajectoriesplotdffirst_2<-separate(data = trajectoriesplotdffirst, col = .id, into = c("exp", "ntrt", "plot"))

trajectoriesplotdfsecond<-plyr::ldply(trajectoriesplotsecond, data.frame)

trajectoriesplotdfsecond_2<-separate(data = trajectoriesplotdfsecond, col = .id, into = c("exp", "ntrt", "plot"))

### add column for decades###

trajectoriesplotdffirst_2<-trajectoriesplotdffirst_2 %>% mutate(decade="1982-1992")

trajectoriesplotdfsecond_2<-trajectoriesplotdfsecond_2 %>% mutate(decade="1993-2004")

##combine them together##
trajectoriesboth<-rbind(trajectoriesplotdffirst_2, trajectoriesplotdfsecond_2)

##mean sinuoisty for each exp_ntrt###

sinuosityavgbydecade<-trajectoriesboth%>% group_by(exp, ntrt, decade) %>% dplyr::summarise(sinuosityavg=mean(sinuosity, na.rm=TRUE),sdsinuosiy=sd(sinuosity, na.rm=TRUE),sesinuosity=sd(sinuosity, na.rm=TRUE)/sqrt(n()), n=n())

```


```{r - recreate FIGURE 4ab Sinuosity}

##for plotting###
sinuosityavg2<-sinuosityavg %>% mutate(ntrt2=ntrt)

## plot specifications
sinuosityavg2$ntrt2  <- mapvalues(sinuosityavg2$ntrt2, from=c("9", "1", "2", "4","6"), to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))
sinuosityavg2$ntrt2 <- factor(sinuosityavg2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


sinuosityavg2$exp<-mapvalues(sinuosityavg2$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

sinuosityavg2$exp<- factor(sinuosityavg2$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))



sinuosityavg2<-sinuosityavg2 %>% mutate(abclabel=exp)
levels(sinuosityavg2$abclabel) <- c( "(a) ", "(b) ")


mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title.x=element_text(size=12))+theme(axis.title.y=element_text(size=10))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))+ theme(panel.grid.major = element_line(colour = "gray 85"))+ theme(panel.grid.minor = element_line(colour = "gray 95"))


sinuosityplot<- ggplot(data=sinuosityavg2, aes(x=ntrt2, y=sinuosityavg, group=ntrt2, color=ntrt2)) +
    geom_point(size=2) +
    geom_errorbar(aes(ymin=lwrconfint, ymax=uprconfint), size=1)+
    facet_grid(~exp)+ 
    xlab("")+
    ylab(expression('Average sinuosity (1982 - 2004)'))+
    mytheme+
    theme(legend.position="none")+
    theme(legend.title = element_blank()) + 
  theme(strip.text.x = element_text(size = 14, colour = "black"))+
   scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
   theme(axis.text.x = element_text(angle = 25,vjust = 0.5))+
    theme(strip.text = element_text(face="bold"))+
     theme(strip.background = element_blank())+
   theme(strip.text = element_text(size=12, face="bold"))+
     theme(strip.background = element_blank())+
      geom_text(aes(x = 0.8, y = 8.2, label = abclabel, group =     abclabel),size = 4, color= "black", check_overlap = T)+ scale_y_continuous(labels = scales::number_format(accuracy = 0.1))+ylim(4.8, 8.3)


```


```{r - recreate SUP FIG 4 Sinuosity 2 decades}



##for plotting###
sinuosityavgbydecade2<-sinuosityavgbydecade %>% mutate(ntrt2=ntrt)

## plot specifications
sinuosityavgbydecade2$ntrt2  <- mapvalues(sinuosityavgbydecade2$ntrt2, from=c("9", "1", "2", "4","6"), to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))

sinuosityavgbydecade2$ntrt2 <- factor(sinuosityavgbydecade2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


sinuosityavgbydecade2$exp<-mapvalues(sinuosityavgbydecade2$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

sinuosityavgbydecade2$exp<- factor(sinuosityavgbydecade2$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))



mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))


mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title.x=element_text(size=12))+theme(axis.title.y=element_text(size=10))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))+ theme(panel.grid.major = element_line(colour = "gray 85"))+ theme(panel.grid.minor = element_line(colour = "gray 95"))


sinuosityplotbydecade<- ggplot(data=sinuosityavgbydecade2, aes(x=ntrt2, y=sinuosityavg, group=ntrt2, color=ntrt2)) +
    geom_point(size=2) +
    geom_errorbar(aes(ymin=sinuosityavg-sesinuosity, ymax=sinuosityavg+sesinuosity), size=1)+
    facet_grid(decade~exp)+ 
    xlab("")+
   ylab(expression('Average sinuosity'))+
    mytheme+
    theme(legend.position="none")+
    theme(legend.title = element_blank()) + 
  theme(strip.text.x = element_text(size = 14, colour = "black"))+
   scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
   theme(axis.text.x = element_text(angle = 25,vjust = 0.5))+
    theme(strip.text = element_text(face="bold"))+
     theme(strip.background = element_blank())+
   theme(strip.text = element_text(size=12, face="bold"))+
     theme(strip.background = element_blank())

jpeg("C:/Users/Melissa/Dropbox/DRAGnet/CedarCreekdata/Dist recovery MS/Figures_overleaf/logfigures/sinuosity2decades.jpeg", width = 6, height = 6, units = 'in', res = 600)
sinuosityplotbydecade
dev.off()

pdf("C:/Users/Melissa/Dropbox/DRAGnet/CedarCreekdata/Dist recovery MS/Figures_overleaf/logfigures/sinuosity2decades.pdf", width = 6, height = 6)
sinuosityplotbydecade
dev.off()




```


```{r - recreate FIG 4 abcd}

g222 <- ggplotGrob(PcOAdecade)
g111 <- ggplotGrob(sinuosityplot)

grid.arrange(g111, g222 , ncol = 1, heights = c(1, 1))



```


##TRAJECTORY DISTANCE####

##
```{r - calculate trajectory distance between years (log data)}

##https://cran.r-project.org/web/packages/ecotraj/vignettes/IntroductionETA.html##

##get the PCoA points into a Euclidean distance matrix to calc. distance between points ###

PCOAandplotsboth_2 <- PCOAandplotsboth %>% arrange(exp, field, ntrt,plot, year)%>%mutate(exp_field_ntrt_plot= paste(exp, field, ntrt,plot, sep = '_'), exp_field_ntrt_plot2 = as.numeric(as.factor(exp_field_ntrt_plot))) %>% dplyr::select(field, exp, plot, subplot, year, disk, ntrt, nadd, expntrtyear, expntrtfieldyear, exp_field_ntrt_plot, exp_field_ntrt_plot2, Axis.1, Axis.2)

###split it by E001 and E002 since there are missing years in E002##
PCOAandplotsE001<-subset(PCOAandplotsboth_2, exp==1) 
PCOAandplotsE002<-subset(PCOAandplotsboth_2, exp==2) 

#drop the years for E002##
levels(PCOAandplotsE002$year)
PCOAandplotsE002$year <- as.factor(as.character(PCOAandplotsE002$year))


EucE001 = dist(PCOAandplotsE001[,13:14])
EucE002 = dist(PCOAandplotsE002[,13:14])


####
trajlengthE001<-trajectoryLengths(EucE001, PCOAandplotsE001$exp_field_ntrt_plot, PCOAandplotsE001$year)

trajlengthE002<-trajectoryLengths(EucE002, PCOAandplotsE002$exp_field_ntrt_plot, PCOAandplotsE002$year)


####
yearnamesE001<-levels(PCOAandplotsE001$year)
trajlengthE001_df=as.data.frame(trajlengthE001)

colnames(trajlengthE001_df)<-yearnamesE001
names(trajlengthE001_df)[names(trajlengthE001_df) == '2004'] <- 'Total'

trajlengthE001_df$names <- rownames(trajlengthE001_df)

trajlengthE001_df2<-trajlengthE001_df %>%separate(names, c("exp", "field", "ntrt", "plot"), "_")

trajlengthE001_df3<-trajlengthE001_df2%>%dplyr::select(exp, field, ntrt, plot, '1982':'Total')

####
yearnamesE002<-levels(PCOAandplotsE002$year)
trajlengthE002_df=as.data.frame(trajlengthE002)
colnames(trajlengthE002_df)<-yearnamesE002
names(trajlengthE002_df)[names(trajlengthE002_df) == '2004'] <- 'Total'

trajlengthE002_df$names <- rownames(trajlengthE002_df)

trajlengthE002_df2<-trajlengthE002_df %>%separate(names, c("exp", "field", "ntrt", "plot"), "_")%>%dplyr::select(exp, field, ntrt, plot, '1982':'Total')

###for this df, we need to drop the columns where we don't have back to back years###
##drop: 1994 (col 17(missing 1995), drop 2000 (missing 2001) (col 21), drop 2002 (missing 2003)(col 22), drop 2003 (missing most of 2004)(col 23)###

trajlengthE002_df3 <- subset(trajlengthE002_df2, select = -c(17,21:23))

##convert wide to long##

traj_longE001 <- gather(trajlengthE001_df3, startingyear, trajdist, '1982':'2003', factor_key=TRUE)

traj_longE002 <- gather(trajlengthE002_df3, startingyear, trajdist, '1982':'1999', factor_key=TRUE)

##combine the two dataframes back together##

traj_longboth<-rbind(traj_longE001, traj_longE002)

##this data frame (traj_long_both) contain the trajectory distance in Euclidean space for each plot between years. traj dist = the dist between starting year and the next year###



##summarize the average distance between years by treatment group##

avgTraj_byyear_E001E002<-traj_longboth%>% group_by(exp,ntrt,startingyear) %>% dplyr::summarise(meantrajdist=mean(trajdist, na.rm=TRUE),sdtrajdist=sd(trajdist, na.rm=TRUE),setrajdist=sd(trajdist, na.rm=TRUE)/sqrt(n()), n=n())

avgTraj_byyear_E001E002_2<- avgTraj_byyear_E001E002 %>% mutate (year2=as.numeric(as.character(startingyear))+0.5-1981)



```


```{r - AIC trajectory analysis, TABLE S5 and S6}


###AIC test whether a linear or nonlinear fit is better##
###linear fit###

trajdistlin<-lm(meantrajdist~year2, data=avgTraj_byyear_E001E002_2)


##nonlinear##

#trajdistnonlin2 <- nls(meantrajdist ~ SSasymp(year2, Asym, R0, lrc), data=avgTraj_byyear_E001E002_2) ##not converging###

trajdistnonlin2_2<- nls(meantrajdist ~ SSmicmen(year2, a, b), data=avgTraj_byyear_E001E002_2)


##simple exponential decay model####

trajdistnonlin3 <- nls(meantrajdist ~ 1 / (1 + year2^c),start = list(c = 1), data=avgTraj_byyear_E001E002_2)


##null##

trajdistnull<-lm(meantrajdist~1, data=avgTraj_byyear_E001E002_2)


rawaic<-AIC(trajdistnonlin3,trajdistnonlin2_2, trajdistlin, trajdistnull)
nR<-dim(avgTraj_byyear_E001E002_2)[1]  #Sample size 
aictable(rawaic,nR)

##linear model with a slope performs better than null and better than nonlinear model###

####linear model equations#####

##subset by experiment##
avgTraj_byyear_E001<-subset(avgTraj_byyear_E001E002_2, exp="Intact in 1982 (E001)")
avgTraj_byyear_E002<-subset(avgTraj_byyear_E001E002_2, exp="Disturbed in 1982 (E002)")


modelsE001 <- dlply(avgTraj_byyear_E001, "ntrt", function(avgTraj_byyear_E001) 
  lm(meantrajdist ~ year2, data = avgTraj_byyear_E001))

# Print the summary of each model
l_ply(modelsE001, summary, .print = TRUE)


modelsE002 <- dlply(avgTraj_byyear_E002, "ntrt", function(avgTraj_byyear_E002) 
  lm(meantrajdist ~ year2, data = avgTraj_byyear_E002))

# Print the summary of each model
l_ply(modelsE002, summary, .print = TRUE)



```


```{r - recreate FIGURE 5ab Trajectory distance w linear fit}

## plot specifications

avgTraj_byyear_E001E002$exp<-mapvalues(avgTraj_byyear_E001E002$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

avgTraj_byyear_E001E002$exp<- factor(avgTraj_byyear_E001E002$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


avgTraj_byyear_E001E002_2<- avgTraj_byyear_E001E002 %>% mutate (year2=as.numeric(as.character(startingyear))+0.5-1981)

avgTraj_byyear_E001E002_2<-avgTraj_byyear_E001E002_2 %>% mutate(ntrt2=ntrt)

avgTraj_byyear_E001E002_2$ntrt2  <- mapvalues(avgTraj_byyear_E001E002_2$ntrt2, from=c("9", "1", "2", "3", "4","5","6"),to=c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))

avgTraj_byyear_E001E002_2$ntrt2 <- factor(avgTraj_byyear_E001E002_2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "2 N + micro", "3.4 N + micro","5.4 N + micro","9 N + micro"))



mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))



###seperate the control from the others...###

avgTraj_byyear_E001E002_2<-avgTraj_byyear_E001E002_2 %>% mutate(abclabel=exp)
levels(avgTraj_byyear_E001E002_2$abclabel) <- c( "(a) ", "(b) ")

avgTrajcontrol<-subset(avgTraj_byyear_E001E002_2, ntrt2=="None")


avgTraj_byyearlinear<- ggplot(data=avgTraj_byyear_E001E002_2) +
    geom_point(data=avgTraj_byyear_E001E002_2, aes(x=as.numeric(as.character(year2)), y=meantrajdist, group=ntrt2, color=ntrt2)) +
    geom_smooth(data=avgTraj_byyear_E001E002_2, aes(x=as.numeric(as.character(year2)), y=meantrajdist, group=ntrt2, color=ntrt2), method = "lm", formula = y ~ x , size = 1, se=FALSE)+
    geom_point(data=avgTrajcontrol, aes(x=as.numeric(as.character(year2)), y=meantrajdist),     color="gray45") +
geom_smooth(data = avgTrajcontrol, aes(x=as.numeric(as.character(year2)), y=meantrajdist), method = "lm", formula = y ~ x, colour = "gray45", size = 0.25)+
    facet_grid(~exp)+ 
    ylab(expression('Annual community trajectory distance'))+
    xlab("Time since experiment (years)")+
    mytheme+
    theme(legend.title = element_blank()) + 
  theme(legend.position = "none") +
    scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
    theme(strip.text = element_text(size=12, face="bold"))+
     theme(strip.background = element_blank())+
 geom_text(aes(x = 2.0, y = 0.25, label = abclabel, group =     abclabel),size = 4, color= "black", check_overlap = T)


##add text and legend##

avgTraj_byyear_2_linear<-avgTraj_byyearlinear+ theme(legend.position = c(0.85, 0.85))+theme(legend.text = element_text(size = 8))+ guides(color = guide_legend(override.aes = list(size = 0.5)))+theme(legend.margin = margin(0, 0, 0, 0))+theme(legend.spacing.y = unit(0, "mm"))+ theme(legend.key.size = unit(1.5, "mm"))



```



## ADDITIONAL SUPPLEMENTAL FIGURES AND ANALYES ###

```{r - Indicator species analyses - 2 time periods log data }

##start w/ exp12subset_2#

##just the plot data##
plot<-exp12subset_1_sorted%>% dplyr::select('field':'ntrt2')

##just the veg data##

veg<-exp12subset_1_sorted%>% dplyr::select(`mass.above.ACHILLEA MILLEFOLIUM(LANULOSA)`:`mass.above.VIOLA SP.`)

###log transformation####

vegmatrix<-as.matrix(veg)

logveg<-log(1+vegmatrix)

##merge back together###

exp12subset_3<-data.frame(plot, logveg)


head(exp12subset_3) ##log transformed values###


###subset to the controls and high N treatments###

exp12subset_4<-subset(exp12subset_3,ntrt==9|ntrt==6)


# Look at first 3 years of experiment and last 3 years of experiment
vegearly <- subset(exp12subset_4, year=="1982"|year=="1983"|year=="1984")
veglate <- subset(exp12subset_4, year=="2000"|year=="2002"|year=="2004")

##skip years 2001, 2003 bc of imcomplete data##


#Put data in long format
vegearlylong <-reshape(vegearly,
                  v.names="mass",
                 timevar="species",
                  idvar=c("field", "exp", "plot", "year", "ntrt", "nadd", "disk"),  
                  times=grep("mass.above", names(vegearly), value=TRUE),
                 varying=list(grep("mass.above", names(vegearly), value=TRUE)),
                  direction="long")

vegearlylong_2<-vegearlylong %>% mutate(stage="early") %>% mutate(stage2=1)


veglatelong <-reshape(veglate,
                  v.names="mass",
                 timevar="species",
                  idvar=c("field", "exp", "plot", "year", "ntrt", "nadd", "disk"),  
                  times=grep("mass.above", names(veglate), value=TRUE),
                 varying=list(grep("mass.above", names(veglate), value=TRUE)),
                  direction="long")

veglatelong_2<-veglatelong %>% mutate(stage="late")%>% mutate(stage2=3)

#####merge the early and late##

veglong2<-rbind(vegearlylong_2, veglatelong_2)


# Delete prefix from species names
veglong2$species <- with(veglong2, gsub("mass.above.","", species))

rownames(veglong2)<-NULL ##get rid of rownames##

##indicator species analysis##

    ##convert data back to wide form##

vegwide2 <- spread(veglong2, species, mass)

vegwideplotdata2<-vegwide2[,1:13]
vegwidematrix2<-vegwide2[,14:224]


###make a column for dist_nitrogen_earylate##

vegwideplotdata22<-vegwideplotdata2 %>% mutate(stagentrtdist= paste(stage, ntrt, exp, sep = '_'))

vegwideplotdata22$stagentrtdist<-as.factor(vegwideplotdata22$stagentrtdist)

vegwideplotdata22$stagentrtdist <- factor(vegwideplotdata22$stagentrtdist, levels = c("early_9_1", "early_6_1", "early_9_2", "early_6_2", "late_9_1", "late_6_1", "late_9_2", "late_6_2"))

vegwideplotdata33<-vegwideplotdata22 %>% mutate(stagentrtdist2= as.numeric(stagentrtdist))

vegwideplotdata33$stagentrtdist2<-as.factor(vegwideplotdata33$stagentrtdist2)

vegwideplotdata33$stagentrtdist2 <- factor(vegwideplotdata33$stagentrtdist2, levels = c("1", "2", "3", "4", "5", "6", "7", "8"))

##early  late ##

##indicator spp analyses###

indvalearlylate = multipatt(vegwidematrix2,vegwideplotdata33$stagentrtdist2,max.order = 3, control = how(nperm=99))## just looks at the individual groups, and twos  ##


indval<-data.frame(indvalearlylate[["sign"]]) ##RESULTS###

```


```{r - SUPPLEMENTAL SPP RICHNESS raw data}


##starting with exp12subset_1###raw biomass data###


###summarize##

###Diversity metrics for plant data###
ShannonD_a <- diversity(exp12subset_1[,11:222], index = "shannon") ##shannon-weiner index##
Richness_a <- specnumber(exp12subset_1[,11:222])
Evenness_a <- ShannonD_a/log(Richness_a)

###combine##
plotdatarichness<-cbind(exp12subset_1[,1:10], Richness_a)

plotdatarichnes2<- plotdatarichness%>%   mutate(ntrt2=ntrt,expntrtyear= paste(exp, ntrt,year, sep = '_'))

##make year numeric##

plotdatarichnes2$year<-as.numeric(as.character(plotdatarichnes2$year))

##average across treatments##

richnessaverage<-plotdatarichnes2%>% group_by(exp, ntrt, ntrt2, year, expntrtyear) %>% dplyr::summarise(meanrichness=mean(Richness_a, na.rm=TRUE),sdrichness=sd(Richness_a, na.rm=TRUE),serichness=sd(Richness_a, na.rm=TRUE)/sqrt(n()), n=n())


##richness at beginning###

richnessbeginning<-subset(plotdatarichnes2, year < 1985)

richnessbegsummary<-richnessbeginning%>% group_by(exp) %>% dplyr::summarise(meanrichness=mean(Richness_a, na.rm=TRUE),sdrichness=sd(Richness_a, na.rm=TRUE),serichness=sd(Richness_a, na.rm=TRUE)/sqrt(n()), n=n())





##richness at end of experiment###
richnessend<-subset(plotdatarichnes2, year > 2000)

richnessendsummary<-richnessend%>% group_by(exp) %>% dplyr::summarise(meanrichness=mean(Richness_a, na.rm=TRUE),sdrichness=sd(Richness_a, na.rm=TRUE),serichness=sd(Richness_a, na.rm=TRUE)/sqrt(n()), n=n())



##plotting specifications###

richnessaverage$exp<-mapvalues(richnessaverage$exp, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

richnessaverage$exp<- factor(richnessaverage$exp, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


richnessaverage_2<- richnessaverage%>% mutate (year2=as.numeric(as.character(year-1981)))


richnessaverage_3<-richnessaverage_2 %>% mutate(ntrt2=ntrt)

richnessaverage_3$ntrt2  <- mapvalues(richnessaverage_3$ntrt2, from=c("9", "1", "2", "4","6"), to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))

richnessaverage_3$ntrt2 <- factor(richnessaverage_3$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))

mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))+ theme(panel.grid.major = element_line(colour = "gray 85"))+ theme(panel.grid.minor = element_line(colour = "gray 95"))




spprichnesstime<-richnessaverage_3 %>% ggplot(aes(x=year2, y=meanrichness, group=ntrt2, color=ntrt2, fill=ntrt2))+
    geom_point()+
    geom_smooth(method="loess")+
    facet_grid(ntrt2~exp) + 
    mytheme+
    ylab("Plot species richness")+
    xlab("Time since experiment")+
    mytheme+
    theme(legend.title = element_blank()) + 
  theme(legend.position = "none") +
  theme(strip.text.x = element_text(size = 14, colour = "black"))+
   scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
   scale_fill_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
     theme(strip.background = element_blank())





```


```{r - SUPPLEMENTAL Biomass over time raw data}

##start with exp12subset_1###

##wide to long form##
exp12subset_long <- gather(exp12subset_1, species, biomass, 'mass.above.ACHILLEA MILLEFOLIUM(LANULOSA)':'mass.above.VIOLA SP.', factor_key=TRUE)

##remove extra text##

exp12subset_long $species<-str_remove(exp12subset_long$species, "mass.above.")


##make a column for exp_ntrt_year##

exp12subset_long_2<-exp12subset_long  %>% mutate( expntrtyear_2= paste(exp,ntrt, year, sep = '_'))

exp12subset_long_3<-exp12subset_long_2 %>% mutate(expntrt= paste(exp,ntrt, sep = '_')) %>% dplyr::select(field, plot, disk,year, ntrt, nadd, expntrtyear_2, expntrt, species, biomass)

##sum all species abundance across plots with the same treatment group and year ###includes zeroes!!!#####

sum_allsp_exp12ABC<-exp12subset_long_3%>% group_by(species, disk,year, ntrt, nadd, expntrtyear_2, expntrt,) %>% dplyr::summarise(sumbiomass=sum(biomass, na.rm=TRUE),n=n())


###get rid of years where we don't have all 18 plots###

sum_allsp_exp12ABC_2<-subset(sum_allsp_exp12ABC, n==18)

sum_allsp_df<-as.data.frame(sum_allsp_exp12ABC_2)


##this dataframe (sumallsp_df has the abundance of each plant in each treatment group (n = 18) for each year####

##calculating the total biomass##

totalbiomassplot<-exp12subset_long_3%>% group_by(plot,field, disk,year, ntrt, nadd, expntrtyear_2, expntrt,) %>% dplyr::summarise(totalbiomass=sum(biomass, na.rm=TRUE),sebiomass=sd(biomass, na.rm=TRUE)/sqrt(n()),n=n())

##average across plots##

totalbiomassgroups<-totalbiomassplot%>% group_by(disk,year, ntrt, nadd, expntrtyear_2, expntrt,) %>% dplyr::summarise(totalplotbiomass=mean(totalbiomass, na.rm=TRUE),seplotbiomass=sd(totalbiomass, na.rm=TRUE)/sqrt(n()), n=n())

totalbiomassgroups2<-subset(totalbiomassgroups, n==18)

totalbiomassdf<-as.data.frame(totalbiomassgroups2)

totalbiomassdf2<-totalbiomassdf%>%mutate(upper=totalplotbiomass+(2*seplotbiomass), lower=totalplotbiomass-(2*seplotbiomass))%>% dplyr::select(expntrtyear_2,year, disk, ntrt, totalplotbiomass, seplotbiomass,upper, lower)

####summary across treatments all years###

biomasssummary<-totalbiomassdf%>% group_by(disk) %>% dplyr::summarise(meanbiomass=mean(totalplotbiomass, na.rm=TRUE),sebiomass=sd(totalplotbiomass, na.rm=TRUE)/sqrt(n()),n=n())


##plot specifications ##


totalbiomassdf2$disk<-mapvalues(totalbiomassdf2$disk, from=c("0", "1"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

totalbiomassdf2$disk<- factor(totalbiomassdf2$disk, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


##
totalbiomassdf22<-totalbiomassdf2 %>% mutate(ntrt2=ntrt)

totalbiomassdf22$ntrt2  <- mapvalues(totalbiomassdf22$ntrt2, from=c("9", "1", "2", "4","6"), to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))

totalbiomassdf22$ntrt2 <- factor(totalbiomassdf22$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


totalbiomassdf3<- totalbiomassdf22 %>% mutate(year2=as.numeric(as.character(year))-1981)

##this dataframe (totalbiomass df3, has the total biomass of all 18 plots per treatment group for all  years###)


###plotting###

mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))+ theme(panel.grid.major = element_line(colour = "gray 85"))+ theme(panel.grid.minor = element_line(colour = "gray 95"))

totalbiomass<-totalbiomassdf3 %>% ggplot(aes(x=year2, y=totalplotbiomass, color=ntrt2))+
    geom_line() + 
    geom_ribbon(aes(ymin=lower, ymax=upper, fill=ntrt2), alpha=0.5, colour = NA) + 
    facet_grid(ntrt2~disk) + 
    mytheme+
    ylab(expression(paste("Total aboveground biomass g / m  "^2)))+
    xlab("Time since experiment")+
    theme(legend.title = element_blank())+
    theme(legend.key.size = unit(0.5, "cm"))+theme(strip.background = element_blank())+theme(strip.text.x = element_text(size = 14, colour = "black"))+
   scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
   scale_fill_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+ theme(legend.position = "none") 
    
```


```{r - SUPPLEMENTAL Functional groups over time raw data}

##start with the sum_all_sp_df and totabiomassdf2 dataframes ####

## and the original Da (all plot and spp info from CC)

##merge the spp with total plot biomass##

sum_allsp_df_2 <- (merge(sum_allsp_df,totalbiomassdf2, by = 'expntrtyear_2', sort=FALSE, all.x=TRUE))

sum_allsp_df_3<-sum_allsp_df_2 %>% mutate(propbiomass=sumbiomass/totalplotbiomass)

### this data set (sum_allsp_df_3 has the sum of all spp in each year in each dist/ntrt treatmet and the proportion of biomass of that group##

###merge the species info with functional group and other information##

specieslookup<-da %>% distinct(species, .keep_all = TRUE)

specieslookup2<-specieslookup %>% dplyr::select(species,functional.group,duration,lifeform, pathway, origin)


mergedfuncgroup <- (merge(specieslookup2,sum_allsp_df_3, by = 'species', sort=FALSE, all.y=TRUE))

mergedfuncgroup_2<-mergedfuncgroup%>% mutate(functional.group2= paste(functional.group, origin, sep = '_'))

####now sum up by functional group + origin ##

sumfunctionalgroup<-mergedfuncgroup_2%>% group_by(disk,year, ntrt, nadd, expntrtyear_2, expntrt,functional.group, totalplotbiomass) %>% dplyr::summarise(sumbiomass=sum(sumbiomass, na.rm=TRUE)) %>% mutate(propbiomass=sumbiomass/totalplotbiomass)

sumfuntionalgroupdf<-as.data.frame(sumfunctionalgroup)

###reordering so that other is last, first grasses, then forbs##

sumfuntionalgroupdf$functional.group<-factor(sumfuntionalgroupdf$functional.group, levels=c("C3", "C4" , "S", "W", "L", "F"))

sumfuntionalgroupdf2<-sumfuntionalgroupdf[complete.cases(sumfuntionalgroupdf), ]

#drop the levels not sampeled
sumfuntionalgroupdf2$functional.group <- as.factor(as.character(sumfuntionalgroupdf2$functional.group))


##plot specifications####

sumfuntionalgroupdf2$disk<-mapvalues(sumfuntionalgroupdf2$disk, from=c("0", "1"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

sumfuntionalgroupdf2$disk<- factor(sumfuntionalgroupdf2$disk, levels = c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


sumfuntionalgroupdf2<-sumfuntionalgroupdf2 %>% mutate(ntrt2=ntrt)

sumfuntionalgroupdf2$ntrt2  <- mapvalues(sumfuntionalgroupdf2$ntrt2, from=c("9", "1", "2", "4","6"), to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))

sumfuntionalgroupdf2$ntrt2 <- factor(sumfuntionalgroupdf2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


sumfuntionalgroupdf2$year<-as.numeric(as.character(sumfuntionalgroupdf2$year))


sumfuntionalgroupdf3<- sumfuntionalgroupdf2 %>% mutate(year2=year-1981)




mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))





propfunctionalgrouptime<-sumfuntionalgroupdf3 %>% ggplot(aes(x=year2, y=propbiomass, group=functional.group))+
    geom_area(position = 'stack', aes(fill=functional.group)) + 
    facet_grid(ntrt2~disk) + 
    mytheme+
    ylab("Proportion of Biomass")+
    xlab("Time since experiment")+
    theme(legend.title = element_blank())+
    theme(legend.key.size = unit(0.5, "cm"))+scale_fill_viridis_d(option="C", na.value="grey72")+
    theme(strip.background = element_blank())+theme(strip.text.x = element_text(size = 14, colour = "black"))+
    scale_y_continuous(breaks=c(0.25,0.50,0.75, 1.00))
    

```




##additional plots not included in paper###


```{r - pCOA species last few years}


##start with log transformed data###

exp12subset_2004<-subset(exp12subset_2, year==2004)


##just the veg data##

veg2004<-exp12subset_2004%>% dplyr::select(`mass.above.ACHILLEA MILLEFOLIUM(LANULOSA)`:`mass.above.VIOLA SP.`)

#remove columns with all 0s##

veg2004_2 = veg2004[,colSums(veg2004) > 0]

##just the plot data##
plot2004<-exp12subset_2004%>% dplyr::select('field':'ntrt2')%>% mutate(expntrt= paste(exp,ntrt,sep = '_')) 

## Bray curtis dissimiliarty matrix##
dist2004<-vegdist(veg2004_2, method="bray")

##PCOA####
PCOA2004<-pcoa(dist2004)

##plot the ord##

#pcoa1<-plot(PCOA2004)


## extracting the axis scores##

PCOA2004 <- PCOA2004$vectors[,c(1,2)]


##merge the axis scores with the plot data ###

PCOAandplots2004<-cbind(plot2004,PCOA2004)


###pulling out the group centroids ####


bd_2004<-betadisper(vegdist(veg2004_2),plot2004$expntrt, type = "centroid")    ####Note that you have to specify the centroid ##  ##takes a bit##


centroiddf2004<-as.data.frame(bd_2004[["centroids"]])

centroiddf2004_2<-centroiddf2004[,1:2]

centroiddf2004_2<- tibble::rownames_to_column(centroiddf2004_2, "expntrt")

centroiddf2004_2<-separate(data = centroiddf2004_2, col = expntrt, into = c("exp", "ntrt"))




##plotting specifications#####


PCOAandplots2004<-PCOAandplots2004%>% mutate(ntrt2=ntrt)%>% mutate(exp2=exp)

PCOAandplots2004$ntrt2  <- mapvalues(PCOAandplots2004$ntrt2, from=c("9", "1", "2", "4","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))
PCOAandplots2004$ntrt2 <- factor(PCOAandplots2004$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


PCOAandplots2004$exp2<-mapvalues(PCOAandplots2004$exp2, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

PCOAandplots2004$exp2 <- factor(PCOAandplots2004$exp2, levels =c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

####
centroiddf2004_2<-centroiddf2004_2%>% mutate(ntrt2=ntrt)%>% mutate(exp2=exp)

centroiddf2004_2$ntrt2  <- mapvalues(centroiddf2004_2$ntrt2, from=c("9", "1", "2", "4","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))
centroiddf2004_2$ntrt2 <- factor(centroiddf2004_2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


centroiddf2004_2$exp2<-mapvalues(centroiddf2004_2$exp2, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

centroiddf2004_2$exp2 <- factor(centroiddf2004_2$exp2, levels =c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))


mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))




####

PCOA2004plot <- ggplot() + 
    xlab("PCOA 1") +
    ylab("PCOA 2") +  
    geom_point(data=PCOAandplots2004, 
               aes(x=Axis.1, y=Axis.2, colour=ntrt2, shape=exp2), 
               size=1)+
        geom_point(data=centroiddf2004_2, 
               aes(x=PCoA1, y=PCoA2, colour=ntrt2, shape=exp2), 
               size=4)+
  scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+
  scale_fill_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+ 
   mytheme+
    theme(strip.text.x = element_text(size = 14, colour = "black"))+
    theme(strip.text.y = element_text(size = 14, colour = "black"))+
    theme(legend.position = c(0.89, 0.35),
          legend.spacing.x = unit(0.03, 'cm'))+
    theme(legend.text = element_text(size=6))+
    theme(legend.key.size = unit(0.4, 'cm'))


```



```{r - NMDS species last few years}


##start with log transformed data###

exp12subset_2004<-subset(exp12subset_2, year==2004)



##just the veg data##

veg2004<-exp12subset_2004%>% dplyr::select(`mass.above.ACHILLEA MILLEFOLIUM(LANULOSA)`:`mass.above.VIOLA SP.`)

#remove columns with all 0s##

veg2004_2 = veg2004[,colSums(veg2004) > 0]

##just the plot data##
plot2004<-exp12subset_2004%>% dplyr::select('field':'ntrt2')%>% mutate(expntrt= paste(exp,ntrt,sep = '_')) 

## Bray curtis dissimiliarty matrix##
dist2004<-vegdist(veg2004_2, method="bray")



###NMDS###

NMDS1<-metaMDS(dist2004, k = 3, trymax = 10000)

##extract the spp scores##

en = envfit(NMDS1, veg2004_2, permutations = 999, na.rm = TRUE)

endf <- data.frame((en$vectors)$arrows, (en$vectors)$r, (en$vectors)$pvals)

endf <- tibble::rownames_to_column(endf , "species")

colnames(endf ) <- c("species", "NMDS1", "NMDS2", "r2", "pval")

endf$species<-str_remove(endf$species, "mass.above.")


##select out certain spp only####

endfmain<-endf %>% filter(r2 > 0.3)

species2<-c("Agro.rep", "Ambro.cor", "Carex", "Euph.cor", "Lath.ven", "Poa.pra", "Schiz.scop", "Sorg.nut")

enddfmain2<-as.data.frame(cbind(endfmain, species2))

## extracting the axis scores##

NMDSscores<-NMDS1$points

##merge the axis scores with the plot data ###

NMDSandplots2004<-cbind(plot2004,NMDSscores)


###AVERAGE BY EXP_NTRT####

NMDSavg<-NMDSandplots2004 %>% group_by(exp, ntrt) %>% dplyr::summarise(meanAxis1=mean(MDS1, na.rm=TRUE),sdAxis1=sd(MDS1, na.rm=TRUE),seAxis1=sd(MDS1, na.rm=TRUE)/sqrt(n()),meanAxis2=mean(MDS2, na.rm=TRUE),sdAxis2=sd(MDS2, na.rm=TRUE),seAxis2=sd(MDS2, na.rm=TRUE)/sqrt(n()), n=n())



##plotting specifications#####


NMDSandplots2004_2<-NMDSandplots2004%>% mutate(ntrt2=ntrt)%>% mutate(exp2=exp)

NMDSandplots2004_2$ntrt2  <- mapvalues(NMDSandplots2004_2$ntrt2, from=c("9", "1", "2", "4","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))
NMDSandplots2004_2$ntrt2 <- factor(NMDSandplots2004_2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


NMDSandplots2004_2$exp2<-mapvalues(NMDSandplots2004_2$exp2, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

NMDSandplots2004_2$exp2 <- factor(NMDSandplots2004_2$exp2, levels =c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

##plotting specifications#####


NMDSavg_2<-NMDSavg%>% mutate(ntrt2=ntrt)%>% mutate(exp2=exp)

NMDSavg_2$ntrt2  <- mapvalues(NMDSavg_2$ntrt2, from=c("9", "1", "2", "4","6"),
                               to=c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))
NMDSavg_2$ntrt2 <- factor(NMDSavg_2$ntrt2, levels = c("None", "0 N+micro", "1 N + micro", "3.4 N + micro","9 N + micro"))


NMDSavg_2$exp2<-mapvalues(NMDSavg_2$exp2, from=c("1", "2"), to=c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

NMDSavg_2$exp2 <- factor(NMDSavg_2$exp2, levels =c("Intact in 1982 (E001)", "Disturbed in 1982 (E002)"))

####

####


mytheme<- theme_bw()+ theme(axis.line.x= element_line(colour = "black", size=0.3))+theme(axis.line.y= element_line(colour = "black", size=0.3))+theme(axis.text.x=element_text(size=10, colour = "black"))+theme(axis.text.y=element_text(size=10, colour = "black"))+theme(axis.title=element_text(size=12))+theme(plot.title=element_text(size=7) +theme(plot.title = element_text(hjust = 0.5)))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.title = element_blank())+theme(plot.title = element_text(margin=margin(0,0,5,0)))




####

NMDS2004plot <- ggplot() + 
    xlab("NMDS 1") +
    ylab("NMDS 2") + 
    geom_point(data=NMDSandplots2004_2, 
               aes(x=MDS1, y=MDS2, colour=ntrt2, shape=exp2), 
               size=1)+
  geom_point(data=NMDSavg_2, 
               aes(x=meanAxis1, y=meanAxis2, colour=ntrt2, shape=exp2), 
               size=4)+
  geom_segment(data = enddfmain2,
               aes(x = 0, xend = NMDS1/2, y = 0, yend = NMDS2/2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  geom_text(data = enddfmain2, aes(x = NMDS1/2, y = NMDS2/2, label = species2),
            size = 2.5, position=position_jitter(height=0.03, width=0.03))+
  scale_colour_manual(values = c("gray45", "#E69F00", "#009E73","#0072B2","#CC79A7"))+  
  mytheme+
    theme(strip.text.x = element_text(size = 14, colour = "black"))+
    theme(strip.text.y = element_text(size = 14, colour = "black"))+
    theme(legend.position = c(0.85, 0.25),
          legend.spacing.x = unit(0.03, 'cm'))+
    theme(legend.text = element_text(size=6))+
    theme(legend.key.size = unit(0.4, 'cm'))


```

